\problem{} 
\par Given a finite set of items $I$, where each item $i \in I$ has a positive integer size $s(i) \in \mathbb{Z}^{+}$, and two integer bin capacities $B_1$ and $B_2$. Is there a way to partition the set $I$ into two disjoint subsets $I_1$ and $I_2$ such that every item belongs to exactly one subset, and
\[
\sum_{i\in I_1}s(i)\le B_1
\qquad\text{and}\qquad
\sum_{i\in I_2}s(i)\le B_2?
\]
Prove this problem is NP-complete.

\par \noindent
\textit{Hint: You may use a reduction from the NP-complete problem Subset Sum.}

\solution{
\textbf{1. Membership in NP.}
The problem is in NP. Given $(I_1,I_2)$, we can check that it is a partition of
$I$ and compute both total sizes.

\textbf{2. Known NP-complete problem.}
We reduce from \textsc{Subset Sum}, which is NP-complete.

\textbf{3. Polynomial-time reduction.}
Let $(a_1,a_2,\ldots,a_n,T)$ be a \textsc{Subset Sum} instance. Let
\[
A=\sum_{i=1}^n a_i.
\]
We may assume $T\le A$, since if $T>A$ the \textsc{Subset Sum} instance is
trivially a NO-instance and can be mapped to the fixed NO-instance consisting
of one item of size $3$ and capacities $B_1=B_2=1$.

For each number $a_j$, create one item $i_j$ with size $s(i_j)=a_j$. Set
\[
B_1=T
\qquad\text{and}\qquad
B_2=A-T.
\]
This is a polynomial-time construction.

\textbf{4. Correctness of the reduction.}
If the subset sum instance has a subset $S\subseteq \{1,2,\ldots,n\}$ with
\[
\sum_{j\in S}a_j=T.
\]
Put the corresponding items into $I_1$ and all other items into $I_2$. Then
\[
\sum_{i\in I_1}s(i)=T=B_1
\]
and
\[
\sum_{i\in I_2}s(i)=A-T=B_2.
\]
So the two-bin instance is feasible.

Conversely, suppose the two-bin instance has a feasible partition $(I_1,I_2)$.
Since every item is placed in exactly one of the two sets,
\[
\sum_{i\in I_1}s(i)+\sum_{i\in I_2}s(i)=A.
\]
Feasibility gives
\[
\sum_{i\in I_1}s(i)\le T
\qquad\text{and}\qquad
\sum_{i\in I_2}s(i)\le A-T.
\]
If $\sum_{i\in I_1}s(i)<T$, then
\[
\sum_{i\in I_2}s(i)=A-\sum_{i\in I_1}s(i)>A-T,
\]
contradicting feasibility. Therefore
\[
\sum_{i\in I_1}s(i)=T.
\]
Thus the numbers corresponding to $I_1$ form a subset whose sum is exactly
$T$.

Therefore the subset sum instance is a YES-instance iff the constructed
partition instance is a YES-instance. Hence the problem is NP-hard. Since it is
also in NP, it is NP-complete.
}


\newpage
